Concurrent Programming with Harmony

### 2020-07-30 book/paper.tex
Latest

\begin{figure}
\begin{code}
\harmonysource{RWsbs}
\end{code}
\caption{\harmonylink{code/RWsbs.hny} Reader/Writer Lock implementation using Split Binary Semaphores.}
\label{fig:RWsplitsema}
\end{figure}

@@

We have seen solutions to the reader/writer problem and the producer/consumer
problem that do not busy-wait, but they are quite different.  It would be nice
to have a general technique for all CCS problems.
The ``Split Binary Semaphore'' (SBS) is such a technique, originally proposed by
Tony Hoare and popularized by Edsger Dijkstra~\cite{EWD703}.
SBS is an example of what one could call a ``baton-passing'' technique.
A process needs the baton to access shared variables.
When a process has the baton and does not need it any more,
it first checks to see if there are any processes that are waiting for the
baton and can make progress.
If so, it passes the baton directly to one of those processes.
If not, it simply gives up the baton.

@@

Step 1 is to enumerate all waiting conditions.  In the case of the reader/writer
problem, there are two: a process that wants to read may have to wait for a
writer to leave the critical region, while a process that wants to write may
have to wait until all readers have left the critical section.  If there are $N$
waiting conditions, the SBS technique will use $N+1$ binary semaphores: one for
each waiting condition and an extra one that processes use when they first try
to enter the critical section and do not yet know if they need to wait.
We shall call this the \textit{mutex}.

@@

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \textit{r\_entered}: the number of readers in the critical section;
\item \textit{w\_entered}: the number of writers in the critical section;
\item \textit{r\_waiting}: the number of readers waiting to enter;
\item \textit{w\_waiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
To wit, check out \texttt{rlock\_acquire}.  It first checks to see if there
are any writers.  If so, it increments \textit{r\_waiting}.  If not,
it increments \textit{r\_entered}.

>>>>

\begin{figure}
\begin{code}
\harmonysource{RWsbs}
\end{code}
\caption{\harmonylink{code/RWsbs.hny} Reader/Writer Lock using Split Binary Semaphores.}
\label{fig:RWsplitsema}
\end{figure}

@@

We have seen solutions to the reader/writer problem and the producer/consumer
problem that do not busy-wait, but they are quite different.  It would be nice
to have a general technique for all CCS problems.
The ``Split Binary Semaphore'' (SBS) is such a technique, originally proposed by
Tony Hoare and popularized by Edsger Dijkstra~\cite{EWD703}.
SBS is an example of what one could call a ``baton-passing'' technique.
A process needs the baton to access shared variables.
When a process has the baton and does not need it any more,
it first checks to see if there are any processes that are waiting for the
baton and can make progress.
If so, it passes the baton directly to one of those processes.
If not, it releases the baton for some other process to pick up at a
later time.

@@

Step 1 is to enumerate all waiting conditions.  In the case of the reader/writer
problem, there are two: a process that wants to read may have to wait for a
writer to leave the critical region, while a process that wants to write may
have to wait until all readers have left the critical section or until a writer has left.  If there are $N$
waiting conditions, the SBS technique will use $N+1$ binary semaphores: one for
each waiting condition and an extra one that processes use when they first try
to enter the critical section and do not yet know if they need to wait.
We shall call this last semaphore the \textit{mutex}.

@@

Another important part of the SBS technique is to keep careful track of the
state.  In the case of the reader/writer lock, the state consists of the
following shared variables:
\begin{itemize}
\item \textit{r\_entered}: the number of readers in the critical section;
\item \textit{w\_entered}: the number of writers in the critical section;
\item \textit{r\_waiting}: the number of readers waiting to enter;
\item \textit{w\_waiting}: the number of writers waiting to enter.
\end{itemize}
Each of the
\texttt{rlock\_acquire}, \texttt{rlock\_release},
\texttt{wlock\_acquire}, and \texttt{wlock\_release} methods must maintain
this state.
For example, check out \texttt{rlock\_acquire}.
It first checks to see if there
are any writers.  If so, it increments \textit{r\_waiting}.
If not, it increments \textit{r\_entered}.
